---
title: "Politics Are Afoot!"
author: 'w203: Statistics for Data Science'
date: "10/21/2020"
output: 
  pdf_document
---


# The Setup 

There is *a lot* of money that is spent in politics in Presidential election years. So far, estimates have the number at about $11,000,000,000 (11 billion USD). For context, in 2019 Twitter's annual revenue was about \$3,500,000,000 (3.5 billion USD). 

# The work 

Install the package, `fec16`. 

```{r}
## install.packages('fec16')
```

This package is a compendium  of spending and results from the 2016 election cycle. In this dataset are 9 different datasets that cover: 

- `candidates`: candidate attributes, like their name, a unique id of the candidate, the election year under consideration, the office they're running for, etc. 
- `results_house`: race attributes, like the name of the candidates running in the election, a unique id of the candidate, the number of `general_votes` garnered by each candidate, and other information. 
- `campaigns`: financial information for each house & senate campaign. This includes a unique candidate id, the total receipts (how much came in the doors), and total disbursements (the total spent by the campaign), the total contributed by party central committees, and other information. 

# Your task 

Describe the relationship between spending on a candidate's behalf and the votes they receive.

# Your work 

- We want to keep this work *relatively* constrained, which is why we're providing you with data through the `fec16` package. It is possible to gather all the information from current FEC reports, but it would require you to make a series of API calls that would pull us away from the core modeling tasks that we want you to focus on instead. 
- Throughout this assignment, limit yourself to functions that are  within the `tidyverse` family of packages: `dplyr`, `ggplot`, `patchwork`, and `magrittr` for wrangling and exploration and `base`, `stats`, `sandwich` and `lmtest` for modeling and testing. You do not *have* to use these packages; but try to limit yourself to using only these. 

```{r load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(sandwich)
library(lmtest)
library(fec16)
theme_set(theme_minimal())
knitr::opts_chunk$set(dpi = 300, warning = FALSE, message = FALSE) 
```

```{r load data}
candidates    <- fec16::candidates
results_house <- fec16::results_house
campaigns     <- fec16::campaigns

# the candidate ID for Matt Heinz was wrong in the results_house data, as confirmed with the FEC
#  I am changing this.  There are likely many of these, but I came across this one checking the data...
results_house[results_house$cand_id=="H2AZ08078",]$cand_id = "H6AZ02189"
results_house[results_house$cand_id=="H0PA07082",]$general_percent =.594439

```

## 1. What does the distribution of votes and spending look like? 

1. (0 points) In separate histograms, show both the distribution of votes (measured in `results_house$general_percent` for now) and spending (measured in `ttl_disb`). How would you describe what you see in these two plots? Does a log-transform of `ttl_disb` change what you see? 

```{r, message=FALSE, warning=FALSE}
histogram_of_vote <- results_house %>% 
  ggplot(aes(x = general_percent)) + 
  geom_histogram(bins = 200, fill="purple") + 
  labs(
    title = 'Percentage of Vote Won in Campaign', 
    x = 'Vote Share', y = 'Count')
histogram_of_spending <- campaigns %>% 
  ggplot(aes(x = log10(ttl_disb))) + 
  geom_histogram(bins = 150, fill="purple") + 
  labs(title = 'Total Campaign Spending', x = 'Spending $M', y = 'Count') +
  scale_x_continuous(breaks=seq(0, 10, 1),labels = 10^(seq(0,10,1)-6))
histogram_of_vote / histogram_of_spending
```

The vote distribution shows four modes, the greatest being candidates who receive a small (<10%) share of the vote.  There are a few who receive nearly all the vote, and two groups that receive 5-10% more or less than half the votes.

The spending plot shows spending on campaigns mostly in the \$0.1M to \$3M range, with a large peak around $1M.


## 2. How much does a vote "cost"? 

Suppose that you're interested in describing the cost of a vote in the US election. 

2. As a first assessment of vote cost, a colleague suggests that you compute the ratio of `general_votes` divided by `ttl_disb`.  You're skeptical but produce the look so that you can reason about it together. 
    a. (0 points) Join `results_house` and `campaigns` using the `inner_join` function from `dplyr`. (Throughout, we'll refer to this as a `package::function` -- so `dplyr::inner_join`.) 
    b. (0 points) Compute a new variable called `cost_per_vote` that is the ratio of the total general election votes received by a candidate, divided by the total disbursements on that candidate's behalf. For this part of the question, do not log transform `ttl_disb`. 
    c. (0 points) Plot a histogram, using `ggplot` and `geom_histogram` of this `cost_per_vote` variable. If there are outliers on the right tail of the distribution that obscure the distribution, use the `xlim` layer to narrow the window that you show. (But, if you do this, be certain to report that you have dropped observations. Label your plot so that someone can read it clearly.) 

```{r, message=FALSE, warning=FALSE}
d <- inner_join(results_house, campaigns, by = 'cand_id')
d %>% 
  mutate(cost_per_vote = general_votes / ttl_disb) %>%
  ggplot(aes(x = cost_per_vote)) + 
  geom_histogram(bins=100, fill="purple") + 
  xlim(c(-1, 10)) + 
  labs(
    title = '~Cost Per Vote (Actually Votes/$)', 
    x = 'Votes per $', 
    y = 'Count of Observations'
  )
```

I have restricted the data to campaigns that got votes in the general election.  I have done this by removing the rows for which general votes was marked "NA" in the data.
There were data for 1898 campains in total.  Of those, there were finance data for 1342 of them, and of those 880 received votes in the general election.
The difference in 1342 and 880 were candidate who received votes in the primary, but did not participate in the general election.   The total number of campaigns with 
votes in neither the general election and the primary was 4.

The histogram is above, but I find this to be very misleading.  The "Cost per Vote" metric is actually upside-down.
Based on its calculation, it should be labelled Votes/$.  Here is a better graph for Cost Per Vote (log transformed).
Note the peak around $5/vote.  Also, "spending" is a better term than cost, as it reflects association better than "cost" which implies causation.

```{r, message=FALSE, warning=FALSE}
d %>% 
  mutate(Real_cost_per_vote = ttl_disb/general_votes) %>%
  ggplot(aes(x = log10(Real_cost_per_vote ))) + 
  geom_histogram(bins=150, fill="orange") + 
  labs(
    title = 'Spending Per Vote Across Campaigns', 
    x = 'Spending Per Vote ($)', 
    y = 'Number of Campaigns'
  ) +
  #  this is how we label the axis meanginfully for log10 variables
scale_x_continuous(breaks=seq(-4, 10, 1),labels = 10^seq(-4,10,1))  
```


3. (2 points) There seem to be a lot of candidates with a zero cost-per-vote. What might this mean?

The name is bad.  This is votes per dollar...  Zero votes per dollar really means that the candidate got very few votes for the money they spent..  That's what the spike on the left side of the first histogram means.  This either means that they got few votes, or they spent a great deal on few votes, or both.  More on this later.


3b. (0 points) Produce a scatter plot of `general_votes` on the y-axis and `ttl_disb` on the x-axis. What do you note about these distributions? 

```{r, message=FALSE, warning=FALSE}
d %>% 
  ggplot(aes(x = ttl_disb/1000000, y = general_votes/1000)) + 
  geom_point(col="purple") +   
  labs(
    title = 'Relation of Spending to Votes', 
    x = 'Campaign Spending ($M)', 
    y = 'Votes in General Election (K)'
  )
```

There are many campaigns that received very few votes, desite spending millions.  These are on the bottom of the graph.  There are also many campaigns that received many votes despite very little spending.  These are on the left side of the graph.  The remaining campaigns show that most campaigns spent around $1M, but a few campaigns spent more than \$5M.

Note: This is a niaive approach to this question, based on accepted the data as they are in the plot.  A more accurate analysis is included below.

4. (0 points) Now, include information in your plot that identifies whether the candidate doing the spending was a "Democrat", "Republican" or "Other Party". What do you notice about spending patterns? 
  - To do so will require that you `mutate` a new variable for the three-category indicator before plotting. Here's an example of how you might use `mutate` and `case_when` together to create a variable.
  - Once you've produced a variable for candidate party membership, plot your scatter plot again, but this time adding an argument into the `aes()` function that colors the points by party membership. 


Let's also mark which candidates won with a "W" and show incumbents with a larger point.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
d %>% 
  mutate(cost_per_vote = general_votes / ttl_disb) %>%
#  filter(cost_per_vote < 0.5) %>%
#    mutate(sz = .3 + as.integer(incumbent)) %>%
  mutate(WL = ifelse(won==TRUE,"W","")) %>%
    mutate(party = 
           case_when(
             cand_pty_affiliation=="REP" ~ "REP",
             cand_pty_affiliation=="DEM" ~ "DEM",
             TRUE                      ~ "Other")
  ) %>%
  ggplot(aes(x=ttl_disb, y=general_votes)) +
  geom_point(aes(col=party, stroke=as.integer(incumbent)), size=2.5) +
  # make the republic red and democratic blue match the data
    scale_color_manual(values=c("lightblue", "green", "red")) +
    geom_text(aes(label=WL), size=1.5) +
    labs(
    title = 'Relation of Spending to Votes', 
    x = 'Campaign Spending ($)', 
    y = 'Votes in General Election'
  )
```

Another way to present this information is to look at the percentage of votes won versus spending per vote.


```{r, eccho=FALSE, message=FALSE, warning=FALSE}
d2 <- d %>% select(state, district_id, cand_id, primary_percent, general_percent, won, 
                   ttl_disb, general_votes, cand_pty_affiliation, incumbent)

d2 %>%
#  filter(general_percent > .04) %>%
#  filter(general_percent < .9) %>%
  mutate(sz = .3 + as.integer(incumbent)) %>%
  mutate(WL = ifelse(won==TRUE,"W","")) %>%
  mutate(party = 
           case_when(
             cand_pty_affiliation=="REP" ~ "REP",
             cand_pty_affiliation=="DEM" ~ "DEM",
             TRUE                      ~ "Other")
  ) %>%
ggplot(aes(x=log10(ttl_disb/general_votes),y=general_percent)) +
  geom_point(aes( col=party, stroke=as.integer(incumbent)), size=2.5) +
  geom_text(aes(label=WL), size=1.5) +
  scale_color_manual(values=c("lightblue", "green", "red")) +
scale_x_continuous(breaks=seq(-4, 10, 1),labels = 10^seq(-4,10,1)) +
  labs(
    title = 'Percentage of Voting with Spending', 
    x = 'Spending Per Vote ($)', 
    y = 'Percent of Vote Won'
  )

```

The data points in the bottom right part of the graph (and the bottom left of the graph above this one) require some investigation.  They are quite suspicious.  How is it possible to spend $1000 per vote and get such a small percentage of the vote?

After some investigation, these data are mostly for races in CT and NY.  In some states (CT, NY), a single candidate can run under multiple parties, so the candidate can get votes under a number of parties.  This is the story behind those data points in the bottom right.  To restate the problem, in the data, there are some candidates who have multiple rows of data for the $same \hspace{1mm} campaign$ under different parties.  Here is an example from NY:

\tiny
```{r, message=FALSE, warning=FALSE}
nycand <- d %>% 
  filter(state=="NY", district_id=="01", cand_id =="H8NY01148") %>% 
  select(state, district_id, cand_id, cand_name, party, general_votes, general_percent, cand_pty_affiliation, ttl_disb)
nycand
```

\normalsize

Lee Zeldin, a candidate in the First District of New York, who identifies as a Republican, has votes from four party affiliations.  According to the FEC website (https://www.fec.gov/resources/cms-content/documents/partylabels2012.pdf), these are:  

R: Republican, CRV: Conservative, REF: Reform, and IDP: Independence.   

Note that the votes are by party, but the disbursement value is the same for all rows, so the votes (and percentage votes) need to be summed, but the disbursement is value from a single row.  

I group the votes together as a single campaign and use total disbursements to place candidate on these graphs.  The suspicious data at the bottom is gone.

```{r, message=FALSE, warning=FALSE}
# in some states, a candidate can repsresent mutliple parties - group those lines in a single 
#  line for each state/district_id/cand_id
d_single_candidates <- d %>% 
  mutate(primary_votes = 
           ifelse(is.na(primary_votes),0,primary_votes)) %>%  # make primary votes 0 if NA
  mutate(general_votes = 
           ifelse(is.na(general_votes),0,general_votes)) %>%  # make primary votes 0 if NA
  mutate(general_percent = 
           ifelse(is.na(general_percent),0,general_percent)) %>%# make primary votes 0 if NA
  mutate(party2 = 
           case_when(
             cand_pty_affiliation=="REP" ~ "REP",
             cand_pty_affiliation=="DEM" ~ "DEM",
             TRUE                      ~ "Other")
         ) %>%
  mutate(WL = ifelse(won==FALSE,"","W")) %>%
  mutate(INCU = ifelse(incumbent=="TRUE","Incumbent","Challenger")) %>%
  group_by(state, district_id, cand_id) %>%
  summarize(general_votes=sum(general_votes), general_percent=sum(general_percent),
            runoff_votes=sum(runoff_votes), 
            cand_name=first(cand_name),
            primary_votes=sum(primary_votes), 
            cand_pty_affiliation=first(cand_pty_affiliation),
            ttl_receipts=max(ttl_receipts), incumbent=first(incumbent),
            INCU=first(INCU), WL=first(WL), 
            pol_pty_contrib=max(pol_pty_contrib), 
            other_pol_cmte_contrib=max(other_pol_cmte_contrib),
            party2=first(party2),         
            ttl_disb=max(ttl_disb), ttl_indiv_contrib=max(ttl_indiv_contrib)) %>%
  mutate(log_dollars_per_vote = log10((ttl_disb+1)/(general_votes+1))) %>%
  mutate(valid_general = FALSE)


```

Now, let's take a look at the single candidate data, especially those with very low votes and high expeditures.  We need to remove write-in candidates who would have few votes, but whose votes would correspond to higher spending from the primary.  In other words, we will remove data for candidates who lost in the primary and whose general election votes do not reflect the relationship we wish to model.
There are six of these.

```{r}
#  Data wrangling - lot's of it...
#
#  for viewing the runoff loser data
d_runoff_losers <- d_single_candidates %>%
  filter(!is.na(runoff_votes)) %>%
  group_by(state, district_id, cand_pty_affiliation)  %>%
  slice_min(order_by=runoff_votes,n=1) %>%
  ungroup() 

#for use in filtering out these candidates
runoff_losers <- d_runoff_losers %>%
    select(cand_id) %>% unlist

#  six candidates lost in the runoff but got general votes - remove those
d_single_candidates_2 <- d_single_candidates %>% filter(!cand_id %in% runoff_losers)

# now remove candidates that lost multi-party or single-party primaries
# then remove the rest of the candidates that got no general election votes

mp_states <- c("CA","WA")

# mark candidates that lost the primary - in some states this is top in party
#   697 of these
won_primary <- d_single_candidates %>%
  filter(! state %in% mp_states) %>%
  group_by(state, district_id, cand_pty_affiliation) %>%
  slice_max(order_by=primary_votes,n=1) %>%
  ungroup() %>%
  select(cand_id) %>% unlist()

#  mutate(won_party_primary = TRUE, valid_general) %>%
#  select(state, district_id,cand_id, won_party_primary, valid_general) 
  
# in california, etc , top two candidates from any party go to general
# 122 of these
d_won_MP_primary <- d_single_candidates %>%
  filter(state %in% mp_states) %>%
  group_by(state, district_id) %>%
  slice_max(order_by=primary_votes, n=2) %>%
  ungroup() 

# Postma is missing from the DB, so the WA10 race should 
#  not include the 2nd DEM as an MP winner.  Remove her.
won_MP_primary <- d_won_MP_primary %>%
  select(cand_id) %>% 
  filter (cand_id != "H2WA10048") %>%
  unlist 

# should be 818 of these - check
 d_single_candidates_general <- d_single_candidates %>%
   filter(
     ((!(state %in% mp_states)) & (cand_id %in% won_primary)) |
       ((state %in% mp_states) & (cand_id %in% won_MP_primary))) %>%
   # finally, remove the rows associated with no votes in the general election
   #  I have found several of these to be the result of missing candidates in the 
   #   campaign db that caused candidates who were not in the general election
   #  to show up there.
   filter(general_votes > 0)

```
Finally, remove the rows associated with no votes in the general election I have found several of these to be the result of missing candidates in the campaign db that caused candidates who were not in the general election to show up there.

```{r}
d_single_candidates_general %>%
ggplot(aes(x=ttl_disb,y=general_votes)) +
  geom_point(aes( col=party2, stroke=as.integer(incumbent)), size=2) +
  geom_text(aes(label=WL), size=1.5) +
  scale_color_manual(values=c("lightblue", "green", "red")) +
#  ylim(0,10)+
#scale_x_continuous(breaks=seq(0, 3, 1),labels = 10^seq(0,3,1)) +
  labs(
    title = 'Votes Won with Spending', 
    x = 'Spending', 
    y = 'Votes Won'
  )
```


```{r, message=FALSE, warning=FALSE}

d_single_candidates_general %>%
filter(!is.na(general_votes)) %>%
ggplot(aes(x=log10( ttl_disb / general_votes ), 
           y=general_percent)) +
  geom_point(aes( col=party2, stroke=as.integer(incumbent)), size=2) +
  geom_text(aes(label=WL), size=1.5) +
  scale_color_manual(values=c("lightblue", "green", "red")) +
  xlim(-1,6) +
  scale_x_continuous(breaks=seq(0, 6, 1), labels = 10^seq(0,6,1)) +
  labs(
    title = 'Percentage of Voting with Spending' , 
    x = "Spending Per Vote ($)" , 
    y = "Percent of Vote Won")
```

This makes a great deal more sense.

We can also split the data by Winner/Loser and Challenger/Incumbent to look for patterns.

```{r, message=FALSE, warning=FALSE}
d_single_candidates_general %>%
#  filter(won=="TRUE") %>%
#  arrange(desc(total_percent))
  ggplot(aes(x=log_dollars_per_vote, y=general_percent*100)) +
  geom_point(aes(col=party2), size=1.5) +
  scale_color_manual(values=c("blue", "green", "red"))+
  scale_x_continuous(breaks=seq(0, 3, 1),labels = 10^seq(0,3,1)) +

    labs(
    title = 'Percentage of Voting with Spending', 
    x = 'Spending Per Vote ($)', 
    y = 'Percent of Vote Won'
  ) +
#  geom_text(aes(label=WL), size=1.5) + 
  facet_grid(INCU ~WL) +
   theme_bw()
```

It is also informative to look at the winners and the source of their campaign contributions.
If we focus in on Winners for whom individual contributions were low, we can see that their parties contributed funds to give them a fundraising advantage.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
low_alpha <- .4


d_single_candidates_deltas <- d_single_candidates_general %>%   
  group_by(state, district_id) %>%
  summarize(n(),cand_id, general_votes, general_percent, WL, INCU, ttl_disb, ttl_receipts, ttl_indiv_contrib  ,party2, cand_name,pol_pty_contrib, other_pol_cmte_contrib, primary_votes) %>%
  mutate(delta_pct = general_percent - nth(general_percent,2,default=0)) %>%
  mutate(delta_receipts = ttl_receipts - nth(ttl_receipts,2)) %>%
  mutate(delta_disb = ttl_disb - nth(ttl_disb,2,default=0)) %>%
  mutate(delta_indiv_contrib = ttl_indiv_contrib - nth(ttl_indiv_contrib,2,default=0)) %>%
  mutate(delta_pol_pty_contrib = pol_pty_contrib - nth(pol_pty_contrib,2,default=0)) %>%
  mutate(delta_other_pol_cmte_contrib = other_pol_cmte_contrib - nth(other_pol_cmte_contrib,2,default=0)) %>%
  mutate(delta_ttl_pol_contrib = delta_other_pol_cmte_contrib + delta_pol_pty_contrib) %>%
  mutate(delta_votes = general_votes - nth(general_votes,2,default=0)) %>%
  mutate(per_vote_receipts_advantage = delta_receipts/general_votes) %>%
  mutate(per_vote_spending_advantage = delta_disb/general_votes) %>%
  mutate(per_vote_indiv_contrib_advantage = delta_indiv_contrib/general_votes) %>%
  mutate(per_vote_party_contrib_advantage = delta_pol_pty_contrib/general_votes) %>%
  mutate(per_vote_party_advantage = delta_ttl_pol_contrib/general_votes) %>%
  mutate(need_help = ifelse(delta_indiv_contrib/general_votes < 0,1,low_alpha) ) %>%
  mutate(got_help = ifelse(delta_indiv_contrib/general_votes < 0,low_alpha,1) )

just_winners <- d_single_candidates_general %>%   
  group_by(state, district_id) %>%
  summarize(n(),cand_id, general_votes, general_percent, WL, INCU, ttl_disb, ttl_receipts, ttl_indiv_contrib  ,party2, cand_name,pol_pty_contrib, other_pol_cmte_contrib) %>%
  mutate(delta_pct = general_percent - nth(general_percent,2,default=0)) %>%
  mutate(delta_receipts = ttl_receipts - nth(ttl_receipts,2)) %>%
  mutate(delta_disb = ttl_disb - nth(ttl_disb,2,default=0)) %>%
  mutate(delta_indiv_contrib = ttl_indiv_contrib - nth(ttl_indiv_contrib,2,default=0)) %>%
  mutate(delta_pol_pty_contrib = pol_pty_contrib - nth(pol_pty_contrib,2,default=0)) %>%
  mutate(delta_other_pol_cmte_contrib = other_pol_cmte_contrib - nth(other_pol_cmte_contrib,2,default=0)) %>%
  mutate(delta_ttl_pol_contrib = delta_other_pol_cmte_contrib + delta_pol_pty_contrib) %>%
  mutate(delta_votes = general_votes - nth(general_votes,2,default=0)) %>%
  mutate(per_vote_receipts_advantage = delta_receipts/general_votes) %>%
  mutate(per_vote_spending_advantage = delta_disb/general_votes) %>%
  mutate(per_vote_indiv_contrib_advantage = delta_indiv_contrib/general_votes) %>%
  mutate(per_vote_party_contrib_advantage = delta_pol_pty_contrib/general_votes) %>%
  mutate(per_vote_party_advantage = delta_ttl_pol_contrib/general_votes) %>%
  mutate(need_help = ifelse(delta_indiv_contrib/general_votes < 0,1,low_alpha) ) %>%
  mutate(got_help = ifelse(delta_indiv_contrib/general_votes < 0,low_alpha,1) ) %>%
  filter(WL=="W")

just_winners_needing_help <- just_winners %>%  
  filter(per_vote_indiv_contrib_advantage <0)
    
plt0 <- just_winners %>%
  filter(not(delta_votes == 0)) %>%
  ggplot(aes(x=per_vote_receipts_advantage, y=delta_pct, col=INCU)) +
#  ggplot(aes(x=delta_disb, y=delta_votes, col=INCU)) +
  geom_point() +
  xlim(-5,25)+
  xlab("Winner's Receipts Advantage ($/vote)")+
  ylab("Wining Margin") +
  theme(axis.title = element_blank())

plt1 <- just_winners %>%
  filter(not(delta_votes == 0)) %>%
  ggplot(aes(x=per_vote_spending_advantage, y=delta_pct, col=INCU)) +
  geom_point() +
  xlim(-5,25)+
  xlab("Winner's Spending Advantage ($/vote)")+
  ylab("Wining Margin")
#  theme(axis.title.x = element_blank())+
#  theme(legend.position="none")


plt2 <- just_winners %>%
  filter(not(delta_votes == 0)) %>%
  ggplot(aes(x=per_vote_indiv_contrib_advantage, y=delta_pct, col=INCU, alpha=I(need_help))) +
#  ggplot(aes(x=delta_disb, y=delta_votes, col=INCU)) +
  geom_point() +
  xlim(-5,25)+
#  xlab("Winner's Individidual Contribution Advantage ($/vote)")+
  ylab("Wining Margin")+
  theme(axis.title.x = element_blank(),
       legend.position = c(1, 1), 
       legend.justification = c(1, 1))

d_x <- 0.16
plt2b <- just_winners %>%
  filter(not(delta_votes == 0)) %>%
  filter(per_vote_indiv_contrib_advantage <0) %>%
  ggplot(aes( y=delta_pct, col=INCU)) +
#  ggplot(aes(x=delta_disb, y=delta_votes, col=INCU)) +
  geom_segment(aes(x=per_vote_indiv_contrib_advantage, 
                   xend=(delta_pol_pty_contrib + delta_other_pol_cmte_contrib)/general_votes-d_x,
                   yend=delta_pct),
               size=0.3,col="grey50",
                   arrow = arrow(length = unit(0.2, "npc"),type="closed")) +
    geom_point(aes(x=per_vote_indiv_contrib_advantage)) +
  geom_point(aes(x=(delta_pol_pty_contrib + delta_other_pol_cmte_contrib)/general_votes)) +
  xlim(-5,25)+
  ylim(0,.22)+
  scale_y_continuous(breaks=seq(0, .2, .2))+
  ylab("Wining Margin")+
  theme(axis.title.x = element_blank())+
  theme(legend.position="none")

plt2c <- just_winners %>%
  filter(not(delta_votes == 0)) %>%
  ggplot(aes(x=per_vote_indiv_contrib_advantage, y=delta_pct, col=INCU, alpha=I(need_help))) +
  geom_point() +
  geom_point(aes(x=(delta_pol_pty_contrib + delta_other_pol_cmte_contrib)/general_votes)) +
  geom_segment(data=just_winners_needing_help,aes(x=per_vote_indiv_contrib_advantage, 
                   xend=(delta_pol_pty_contrib + delta_other_pol_cmte_contrib)/general_votes-d_x,
                   yend=delta_pct),
               size=0.3,col="grey20",
               arrow = arrow(length = unit(0.02, "npc"),type="closed")) +
  xlim(-5,25)+
  ggtitle("Party's Contribute to Provide an Advantage for Candidates \n When Individual Contributions are Low ($/vote)")+
  xlab("Winner's Spending Advantage ($/vote)")+
  ylab("Wining Margin")+
  theme(legend.position = c(1, 1), 
       legend.justification = c(1, 1))

plt3 <- just_winners %>%
    filter(not(delta_votes == 0)) %>%
#  filter(not(delta_pol_pty_contrib == 0)) %>%
  ggplot(aes(y=delta_pct)) +
  geom_point(aes(x=per_vote_spending_advantage -  per_vote_party_advantage , col=INCU)) +
  #geom_point(aes(x=log10(pol_pty_contrib)), col="grey20") +
  #geom_point(aes(x=log10(delta_pol_pty_contrib)), col="grey70") +
  #geom_point(aes(x=log10(delta_disb) , col=INCU))+
  xlim(-5,25)+
  xlab("Winner's Non-Party Spending ($/vote)")+
  ylab("Wining Margin")


plt4 <- just_winners %>%
  filter(not(delta_votes == 0)) %>%
  #filter(not(delta_pol_pty_contrib == 0)) %>%
  ggplot(aes(y=delta_pct)) +
  geom_point(aes(x=(delta_pol_pty_contrib + delta_other_pol_cmte_contrib)/general_votes, col=INCU)) +
  #geom_point(aes(x=log10(pol_pty_contrib)), col="grey20") +
  #geom_point(aes(x=log10(delta_pol_pty_contrib)), col="grey70") +
  #geom_point(aes(x=log10(delta_disb) , col=INCU))+
  xlim(-5,25)+
  xlab("Winner's Party's Contribution Advantage ($/vote)")+
  ylab("Wining Margin")


plt5 <- just_winners %>%
  filter(not(delta_votes == 0)) %>%
  #filter(not(delta_pol_pty_contrib == 0)) %>%
  ggplot(aes(y=delta_pct)) +
  geom_point(aes(x=(per_vote_spending_advantage - delta_pol_pty_contrib + delta_other_pol_cmte_contrib)/general_votes, col=INCU)) +
  #geom_point(aes(x=log10(pol_pty_contrib)), col="grey20") +
  #geom_point(aes(x=log10(delta_pol_pty_contrib)), col="grey70") +
  #geom_point(aes(x=log10(delta_disb) , col=INCU))+
  xlim(-5,25)+
  xlab("Winner's Non-Party Advantage ($/vote)")+
  ylab("Wining Margin")

#library(gridExtra)
#grid.arrange(plt2, plt2b, plt1, ncol=1)
#plot_grid(plt2, plt2b, plt1, align = "v", nrow = 3, rel_heights = c(0.41, 0.18, 0.41))
plt2c
```


We can also look at the spending advantage of all winning candidates.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
plt1
```
```{r}
paste(100 - round(sum(just_winners$per_vote_spending_advantage<0) / 
                    length(just_winners$per_vote_spending_advantage)*100,2), 
      "percent of winners outspend their rivals. ")
```


```{r}
t.test(just_winners$per_vote_spending_advantage[just_winners$INCU=="Incumbent"])
```

```{r}
t.test(just_winners$per_vote_spending_advantage[just_winners$INCU=="Challenger"])
```

From the t-test, we can see that winning Incumbents outspend their opponents by \$5.64 per vote, 95% CI [\$5.01 - \$6.27].
Similarly, winning challengers outspend their opponents but by a lower margin: \$2.79 per vote, 95% CI [\$1.07 - \$4.50].


```{r}
pctt <- 100 - round(sum(just_winners$INCU=="Challenger")/length(just_winners$INCU) *100,2)
paste(pctt, "percent of Incumbents win. ")
```


# Produce a Descriptive Model 

5. (10 Points) 

- Given what you have now observed in the data, produce a descriptive model that you think does a good job at describing the relationship between candidate spending and votes they receive. In this model, please include at least three model features and interpret any coefficients you estimate. 

- Tasks to keep in mind as you're writing about your model: 
    - At the time that you're writing and interpreting your regression coefficients you'll be *deep* in the analysis. Nobody will know more about the data than you do, at that point. *So, although it will feel tedious, be descriptive and thourough in what you think you're observing in your model.* 
    - It can be hard to strike the balance between: on the one hand,  writing enough of the technical underpinnings to know that your model meets the assumptions that it must; and, on the other hand, writing little enough about the model assumptions that the implications of the model can still be clear. We're starting this practice now, so that by the end of Lab 2 you will have had several chances to strike this balance.
    

Given what I have observed:

1. 98.4% of winners outspend their closet rival.
2. 86.7% of incumbents win.
3. Political parties of winners make up for a disadvantage in individual constributions with party contributions.
4. Winning incumbents outspend their opponents by \$5.64 per vote, 95% CI [\$5.01 - \$6.27].
5. Winning challengers outspend their opponents but by a lower margin: \$2.79 per vote, 95% CI [\$1.07 - \$4.50].


MODEL 1

$$ Votes = \beta_0 + \beta_1\cdot Incumbent + \beta_2\cdot \mbox {log(Total disbursements)} + \beta_3 \cdot party $$

$\beta_0$ will be the percentage of votes that a non incumbent with no spending advantage gets.
$\beta_1$ will be the incremental votes that an incumbent gets.
$\beta_2$ will be the incremental votes resulting increasing disbursements by 1%
$\beta_3$ will be the incremental votes resulting from being associated with a party.

```{r}
d_single_candidates$INCU <- as.factor(d_single_candidates$INCU)

model1 <- lm(general_votes ~ INCU + log(ttl_disb+1) + party2, data=d_single_candidates_general)
summary(model1)

```

Model 1 does a pretty good job of explaining the varianes in General Election votes by campaign.
Incumbency is worth an additional 54,850 votes, while being in Indepent costs a candidate about 55,000 votes as compared with a Democrat.
Increasing spending by 1% provides an additional 9712 votes on average.  The difference in Party benefit between Republicans and Democrats is not statistically significant.


MODEL 2
$$ Votes = \beta_0 + \beta_1\cdot Incumbent + \beta_2\cdot \mbox {log(Total disbursements)} + \beta_3 \cdot party + \beta_4 \cdot State$$

$\beta_0$ will be the percentage of votes that a non incumbent with no spending advantage gets.
$\beta_1$ will be the incremental votes that an incumbent gets.
$\beta_2$ will be the incremental votes resulting increasing disbursements by 1%
$\beta_3$ will be the incremental votes resulting from being associated with a party.
$\beta_4$ will be the incremental votes resulting from the election being in a specific State.

```{r}
model2 <- lm(general_votes ~ INCU + log(ttl_disb+1) + party2 + state, data=d_single_candidates_general)
summary(model2)


```

Model 2 does an even better job of explaining the varianes in General Election votes by campaign.
The coefficients on Incumbency, and Disbursements are similar to Model 1 (Party=I is -80K).  The coefficients on States are mostly not statistically significant, but jointly, they improve adjusted R-square to nearly 70%!  

I interpret this as is illustrated in the graphs above.  What is important is not just to spend alot, but to outspend your opponent.  By doing per-state analysis, we are essentially comparing spending among candidates within each state, making their spending relative to each other more important.   It can also reflect the percentage voter turnout in the state, and thus the total number of votes.  Further each state has its own rules for primaries and runnoffs, so these differences may be captured in the State coefficients.

I chose not to model spending relative to the opponent, due to causality concerns.  That is, outspending requires knowledge of the opponent's spending.  It is possible to know this until late in the campaign based on FEC reporting requirements, but there is likely large amounts of spending very late in the campaign that is not based on knowing the final spending by the opponent, as it would not be possible to know this.
